---
title: "MAT1856/APM466 Assignment 1"
author: "Xin Wei, 1003058594"
date: "Feburary, 2020"
output: pdf_document
---

\section{Fundamental Questions - 25 points}
\begin{enumerate}
    \item \hfill
    \begin{enumerate}
        \item A government issues bond to raise funds to support government spending.
        \item The yield curve demonstrates the yields (interest rates) of government
        bonds at different maturities, and helps governments analyze 
        economic activity.
        \item The government can sell bonds to reduce money supply by removing 
        cash from the economy in exchange for bonds.
    \end{enumerate}
    \item To begin with, I will eliminate long-term bonds (i.e., at least 30 years),
    including CAN 10.50 Mar 15, CAN 9.75 Jun 1, CAN 9.25 Jun 1, CAN 9.00 Jun 1, CAN
    8.00 Jun 1, CAN 5.75 Jun 1 and CAN 8.00 Jun 1. As noticed, these bonds have high
    coupon rates due to long time to maturity and the illiquidity. If they are
    included in the calculation, they will put more weights on earlier spot rates and
    thus the geometric mean of spot rate, YTM, will slightly variate. Apart from that,
    I will omit on-the-run bonds that are maturing in over 5 years. Since we only need
    those maturing in 0-5 years to draw the curve. Hence, CAN 1.25 Mar 1, CAN 1.00 Jun
    1, CAN 2.00 Jun 1, 2.25 Jun 1, 1.5 Jun 1 and CAN 2.25 Jun 1 will be safely
    eliminated. After filtration, I will choose from the remaining 19 bonds, in a way
    that bonds are maturing in about every half year. Therefore, I will select CAN 1.5
    Mar 1, CAN 0.75 Sep 1, CAN 0.75 Mar 1, CAN 0.75 Sep 1, CAN 0.5 Mar 1, CAN 2.75 Jun
    1, CAN 1.75 Mar 1, CAN 1.5 Jun 1, CAN 2.25 Mar 1, CAN 1.5 Sep 1. Among these
    bonds, CAN 2.75 Jun 1 and CAN 1.5 Jun 1, maturing in 2.33 and 3.33 years, are
    included because we do not observe any maturing in 2.6 and 3.6 years. 
    \item Since the covariance matrix of a sequence of stochastic processes is
    symmetric, we can use spectral decomposition to expand such matrix in its
    eigenvalues and orthogonal eigenvectors. Each eigenvalue corresponds to each row
    linear combination of points/processes in the covariance matrix. Also, it is
    noticed that eigenvalues decrease exponentially from $\lambda_1$ to $\lambda_n$.
    In other words, if we see the covariances as time series, the future rates and
    yields tend to be eventually stable compared to current ones.

\end{enumerate}




\section*{Empirical Questions - 75 points} 

4. \hfill
     (a) \hfill
```{r, echo=FALSE, message=FALSE}
#cleaning data
library(readxl)
library(dplyr)
bonds_data <- read_xlsx("~/Desktop/bonds_data.xlsx")


bonds_data$issue <- as.POSIXct(bonds_data$`issue date`, format="%m/%d/%Y")
issue <- as.Date(bonds_data$issue)

bonds_data$maturity <- as.POSIXct(bonds_data$`maturity date`, format="%m/%d/%Y")
maturity<- as.Date(bonds_data$maturity)

bonds_data$term <- as.numeric(round((bonds_data$maturity - bonds_data$issue)/365,2))
bonds_data <- arrange(bonds_data, by=term)
bonds_data$ttm <- as.numeric(round((
  bonds_data$maturity - 
    as.POSIXct(Sys.Date(),format="%Y-%m-%d"))/(365),2))
bonds_data <- bonds_data[,-c(3,4)]
bonds_data <- bonds_data[c(1,2,13,14,15,16,12,11,10,9,8,7,6,5,4,3)]

bonds_data <- subset(bonds_data, term <= 30)
bonds_data <- subset(bonds_data, ttm <= 5.5)
#View(bonds_data)
sel_bonds1 <- bonds_data[7:15, ]
sel_bonds2 <- bonds_data[17:18, ]
sel_bonds <- arrange(rbind(sel_bonds1, sel_bonds2), by=ttm)
#View(sel_bonds)



```

```{r, echo=FALSE, message=FALSE}
library(jrvFinance)
prices <- sel_bonds[, 7:16]
#View(prices)
prices_t <- t(prices)
#View(prices_t)
issue <- as.Date(sel_bonds$issue)
maturity <- as.Date(sel_bonds$maturity)
coupon <- sel_bonds$coupon/100

second <- "2020-01-02"
third <- "2020-01-03"
sixth <- "2020-01-06"
seventh <- "2020-01-07"
eighth <- "2020-01-08"
nineth <- "2020-01-09"
tenth <- "2020-01-10"
thirteenth <- "2020-01-13"
fourteenth <- "2020-01-14"
fifteenth <- "2020-01-15"

settle_dates <- c(second, third, sixth, seventh, eighth, nineth, tenth, thirteenth,
                  fourteenth, fifteenth)



ytm1 = vector()
ytm2 = vector()
ytm3 = vector()
ytm4 = vector()
ytm5 = vector()
ytm6 = vector()
ytm7 = vector()
ytm8 = vector()
ytm9 = vector()
ytm10 = vector()
ytm11 = vector()

```

```{r, echo=FALSE, message=FALSE}
i = 1
j = 1
#1
while(i < 12)
{
  if(i == 1)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm1 <- c(ytm1, y)
      j = j + 1
    }
  }
  i = i + 1
  
}

```

```{r, echo=FALSE, message=FALSE}
i = 2
j = 1
#2
while(i < 12)
{
  if(i == 2)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm2 <- c(ytm2, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm2)
```

```{r, echo=FALSE, message=FALSE}
i = 3
j = 1
#3
while(i < 12)
{
  if(i == 3)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm3 <- c(ytm3, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm3)
```
```{r, echo=FALSE, message=FALSE}
i = 4
j = 1
#4
while(i < 12)
{
  if(i == 4)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm4 <- c(ytm4, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm4)
```
```{r, echo=FALSE, message=FALSE}
i = 5
j = 1
#5
while(i < 12)
{
  if(i == 5)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm5 <- c(ytm5, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm5)
```
```{r, echo=FALSE, message=FALSE}
i = 6
j = 1
#6
while(i < 12)
{
  if(i == 6)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm6 <- c(ytm6, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm6)
```
```{r, echo=FALSE, message=FALSE}
i = 7
j = 1
#7
while(i < 12)
{
  if(i == 7)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm7 <- c(ytm7, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm7)
```
```{r, echo=FALSE, message=FALSE}
i = 8
j = 1
#8
while(i < 12)
{
  if(i == 8)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm8 <- c(ytm8, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm8)
```
```{r, echo=FALSE, message=FALSE}
i = 9
j = 1
#9
while(i < 12)
{
  if(i == 9)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm9 <- c(ytm9, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm9)
```
```{r, echo=FALSE, message=FALSE}
i = 10
j = 1
#10
while(i < 12)
{
  if(i == 10)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm10 <- c(ytm10, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm10)


```


```{r, echo=FALSE, message=FALSE}
i = 11
j = 1
#10
while(i < 12)
{
  if(i == 11)
  {
     while(j < 11)
    {
      y <- bond.yield(settle=settle_dates[j], mature=maturity[i], coupon=coupon[i], 
                      price=prices_t[j,i])
      ytm11 <- c(ytm11, y)
      j = j + 1
    }
  }
  i = i + 1
  
}
#print(ytm11)


```






```{r, echo=FALSE, message=FALSE}
library(lme4)
library(ggplot2)
groups <- c(ytm1, ytm2, ytm3, ytm4,ytm5,ytm6,ytm7,ytm8,ytm9,ytm10)
ytm <- round(rbind(ytm1, ytm2, ytm3, ytm4,ytm5,ytm6,ytm7,ytm8,ytm9,ytm10),4)
ytm <- as.data.frame(ytm)
rownames(ytm) <- c("bond1", "bond2", "bond3", "bond4", "bond5", "bond6", "bond7", "bond8", "bond9", "bond10")
knitr::kable(ytm, caption = "Yields to Maturity")

ttm <- factor(c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5))

ggplot(ytm, aes(x = ttm)) + 
  geom_line(aes(y = ytm[,1], group = 1, color = "Jan2")) +
  geom_line(aes(y = ytm[,2], group = 1, color = "Jan3")) +
  geom_line(aes(y = ytm[,3], group = 1, color = "Jan6")) +
  geom_line(aes(y = ytm[,4], group = 1, color = "Jan7")) +
  geom_line(aes(y = ytm[,5], group = 1, color = "Jan8")) +
  geom_line(aes(y = ytm[,6], group = 1, color = "Jan9")) +
  geom_line(aes(y = ytm[,7], group = 1, color = "Jan10")) +
  geom_line(aes(y = ytm[,8], group = 1, color = "Jan13")) +
  geom_line(aes(y = ytm[,9], group = 1, color = "Jan14")) +
  geom_line(aes(y = ytm[,10], group = 1, color = "Jan15")) +
  labs(x = "Time to Maturity", y = "Yields")


```

\hfill
     (b)\hfill
```{r, eval=FALSE}
spot <- vector()
for(i in 2:5){
  Calculate the spot rates for the first bond, which matures in 0.5 years, 
  using the formula:
    spot rate = log((100+coupon/2)/price)/time_to_maturity
  Then get the lists of time_to_maturity and cashflows of the 2nd, 3rd, 4th, 
  and 5th bonds
  for(j in 1:9)
  {
    Calculate the last discounted cashflow of each bond, saved as "denominator"
  }
  Then spot_rate = log(cashflow/"denominator")/time_to_maturity
}

for(i in 1:5)
{
  Again, calculate the last discounted cashflow of each bond from the 
  6th to the 10th bond and the spot rates
}
```

```{r,echo=FALSE, message=FALSE}
vector<-c(0.009159414, 0.022518716, 0.019293691, 0.013475547, 0.017210570, 0.018071099,0.018346646, 0.019263979, 0.031018013, 0.015205762, 0.012476315, 0.008695744,0.022307578, 0.018989201, 0.013375744, 0.016945541, 0.017792818, 0.017998186,0.018898096, 0.030877646, 0.014964277, 0.012257877, 0.009170057, 0.022592799,0.018950656, 0.013444151, 0.016867601, 0.017710981, 0.017848476, 0.018740900,0.030861752, 0.014856163, 0.012156323, 0.009339873, 0.022845054, 0.019085269,0.013528847, 0.016937720, 0.017784605, 0.017964895, 0.018863140, 0.030910618,0.014934429, 0.012225143, 0.009516097, 0.022787026, 0.019309325, 0.013490141,0.017008024, 0.017858426, 0.018014873, 0.018915617, 0.030959554, 0.014967012,0.012252925, 0.009699099, 0.022885395, 0.019356483, 0.013513252, 0.017223292,0.018084456, 0.018332006, 0.019248607, 0.031089969, 0.015183018, 0.012445428,0.009889277, 0.022984616, 0.019760818, 0.013536442, 0.017197517, 0.018057393,0.018482739, 0.019406876, 0.031084787, 0.015284671, 0.012535216, 0.010507357,0.023447152, 0.019727194, 0.013668854, 0.017459426, 0.018332397, 0.018668420,0.019601841, 0.031259710, 0.015406467, 0.012640048, 0.010730918, 0.023390243,0.019775958, 0.013629991, 0.017433810, 0.018305500, 0.018551511, 0.019479087,0.031254619, 0.015324411, 0.012564859, 0.010964198, 0.023171816, 0.019553996,0.013528431, 0.017310844, 0.018176386, 0.018333740, 0.019250427, 0.031195020,0.015173260, 0.012427646)

v1<-vector[1:11]
v2<-vector[12:22]
v3<-vector[23:33]
v4<-vector[34:44]
v5<-vector[45:55]
v6<-vector[56:66]
v7<-vector[67:77]
v8<-vector[78:88]
v9<-vector[89:99]
v10<-vector[100:110]
spot_rates<-as.data.frame(rbind(v1, v2,v3,v4,v5,v6,v7,v8,v9,v10))

ggplot(spot_rates, aes(x = ttm)) + 
  geom_line(aes(y = spot_rates[1,], group = 1, color = "Jan2")) +
  geom_line(aes(y = spot_rates[2,], group = 1, color = "Jan3")) +
  geom_line(aes(y = spot_rates[3,], group = 1, color = "Jan6")) +
  geom_line(aes(y = spot_rates[4,], group = 1, color = "Jan7")) +
  geom_line(aes(y = spot_rates[5,], group = 1, color = "Jan8")) +
  geom_line(aes(y = spot_rates[6,], group = 1, color = "Jan9")) +
  geom_line(aes(y = spot_rates[7,], group = 1, color = "Jan10")) +
  geom_line(aes(y = spot_rates[8,], group = 1, color = "Jan13")) +
  geom_line(aes(y = spot_rates[9,], group = 1, color = "Jan14")) +
  geom_line(aes(y = spot_rates[10,], group = 1, color = "Jan15")) +
  labs(x = "Time to Maturity", y = "Yields")


```





5. \hfill
```{r, echo=FALSE, message=FALSE}
five_bonds <- sel_bonds[7:11,]
#View(five_bonds)
fifty_prices <- five_bonds[,7:16]
#View(fifty_prices)
#first_nine_dates <- as.numeric(fifty_prices[,1:9])
#last_nine_dates <- as.numeric(fifty_prices[,2:10])

ln_r1 <- vector()
ln_r2 <- vector()
ln_r3 <- vector()
ln_r4 <- vector()
ln_r5 <- vector()

#bond1
i = 1
while(i < 10)
{
  log_r<-log(fifty_prices[1,i+1] / fifty_prices[1,i])
  ln_r1 = c(ln_r1, log_r)
  i = i + 1
}
ln_r1<-round(as.numeric(ln_r1),7)
#print(ln_r1)
```

```{r, echo=FALSE, message=FALSE}
#bond2
i = 1
while(i < 10)
{
  log_r<-log(fifty_prices[2,i+1] / fifty_prices[2,i])
  ln_r2 = c(ln_r2, log_r)
  i = i + 1
}
ln_r2<-round(as.numeric(ln_r2),7)
```


```{r, echo=FALSE, message=FALSE}
#bond3
i = 1
while(i < 10)
{
  log_r<-log(fifty_prices[3,i+1] / fifty_prices[3,i])
  ln_r3 = c(ln_r3, log_r)
  i = i + 1
}
ln_r3<-round(as.numeric(ln_r3),7)
```

```{r, echo=FALSE, message=FALSE}
#bond4
i = 1
while(i < 10)
{
  log_r<-log(fifty_prices[4,i+1] / fifty_prices[4,i])
  ln_r4 = c(ln_r4, log_r)
  i = i + 1
}
ln_r4 <-round(as.numeric(ln_r4),7)
```

```{r, echo=FALSE, message=FALSE}
#bond5
i = 1
while(i < 10)
{
  log_r<-log(fifty_prices[5,i+1] / fifty_prices[5,i])
  ln_r5 = c(ln_r5, log_r)
  i = i + 1
}
ln_r5<-round(as.numeric(ln_r5),7)
```

```{r, echo=FALSE, message=FALSE}
log_ret <- t(rbind(ln_r1, ln_r2, ln_r3, ln_r4, ln_r5))
colnames(log_ret) <- c("bond1", "bond2", "bond3", "bond4", "bond5")
knitr::kable(log_ret, cap="Matrix of Log-Returns")
cov_lr <- cov(log_ret)
colnames(cov_lr) <- c("bond1", "bond2", "bond3", "bond4", "bond5")
rownames(cov_lr) <- c("bond1", "bond2", "bond3", "bond4", "bond5")
knitr::kable(cov_lr,cap="Covariance Matrix of Log-Returns")

```

6. \hfill
```{r, echo=FALSE, message=FALSE}
eigv_lr <- eigen(cov_lr)
eval_lr <- eigv_lr$values
evec_lr <- eigv_lr$vectors
knitr::kable(eval_lr, cap="Eigenvalues for Covariance Matrix of Log-Return")
knitr::kable(evec_lr, 
             caption = "Eigenvectors for Covariance Matrix of Log-Return")

```

The first eigenvalue weights about 86% of the sum of all eigenvalues from the covariance matrix, so the first pair of eigenvalue and eigenvector well explained the data. 

\section{References and GitHub Link to Code}
1. Boundless. Boundless Finance. Retrieved from https://courses.lumenlearning.com/boundless-finance/chapter/valuing-bonds/
2. Chen, J. (2020, January 29). Yield to Maturity (YTM). Retrieved from https://www.investopedia.com/terms/y/yieldtomaturity.asp
3. Investopedia. (2020, January 29). Yield to Maturity – YTM vs. Spot Rate: What's the Difference? Retrieved from https://www.investopedia.com/ask/answers/020215/what-difference-between-yield-maturity-and-spot-rate.asp
4. Bai, Z. (2010). The Generalized Eigenproblem. G.W. Stewart, 105–109. doi: 10.1007/978-0-8176-4968-5_9
5. Mossberg, M. (2008). Estimation of Continuous-Time Stochastic Signals From Sample Covariances. IEEE Transactions on Signal Processing, 56(2), 821–825. doi: 10.1109/tsp.2007.907829
6. Chen, J. (2020, January 29). The Types of Government Bonds Investors Can Buy. Retrieved from https://www.investopedia.com/terms/g/government-bond.asp

GitHub_link:https://github.com/Xin-Wei-Cynthia/APM466a1/tree/master























